---
title: "HW3 EDA Visualization"
author: "Utkal Sirikonda"
output: html_document
---

# 1. Required Packages for this Analysis

```{r warnings=FALSE, message=FALSE}

library(tidyverse)
library(AER)

help("Guns", package="AER")
data("Guns", package="AER")
```
<br>

# 2. Dataset

I have chosen to work with the Guns dataset from the AER package. 

<br>

## 3. Description: 

Guns is a balanced panel of data on 50 US states, plus the District of Columbia (for a total of 51 states), by year for 1977–1999.

<br>


## 4. First 10 Rows

The first 10 rows of the dataset look like the following. 

```{r}
head(Guns, 10)
```

<br>

## 5. Number of Observations and Variables

```{r}
nrow(Guns)

ncol(Guns)
```

<br>

## 6. View of types of data

Below is a visual of the types of data in my dataset. 

```{r}
# CODE TO GENERATE VIEW OF DATA TYPES
visdat::vis_dat(Guns)
```

<br>

# Numeric Variables

## 7. Basic Statistcs 

Below are the basic descriptive statistics of the "prisoners" variable.

```{r}
summary(Guns$prisoners)

mean(Guns$prisoners)
sd(Guns$prisoners)
min(Guns$prisoners)
max(Guns$prisoners)
```

<br>

## 8. Graph of one variable

```{r}
ggplot(Guns, aes(x = violent)) +
  geom_histogram(binwidth = 50) +
  labs(title = "Distribution of Violent Crime Rates", x = "Violent Crime Rate", y = "Frequency")
```

<br>

## 9. Description of Graph

The graph is left skewed. We can observe that majority of crimes are less violent and more violent crime rates are less frequent.

<br>

# Relationship between variables.

Below is an analysis of how Violence depends on Density. 


<br>

## 10. Graph of two variables

```{r}
ggplot(Guns, aes(x = violent, y = density)) +
  geom_line() + geom_point() +
  labs(title = "Scatter plot of Violent Crime Rate vs Density",
       x = "Density of Population",
       y = "Violent Crime Rate (violent)")
```

<br>

## 11. Description of Graph

The Violent Crime Rate is high where there is more dense population  








