---
title: "Merging Lab Results"
author: "Utkal Sirikonda"
output: html_document
---

### Setting Up

Packages used in this lab. 

```{r warning=FALSE, message=FALSE}
library(tidyverse)
```


```{r include=FALSE}
# Leave this r chunck AS IS

enrollment <- read_csv("enrollment.csv")
survey <- read_csv("survey.csv")

purchases <- read_csv("purchases.csv")
prices <- read_csv("prices.csv")
```




## Case Study 1: Student Records

### Joining the `enrollment` and `survey` data so that you have a list of the students in the class with their survey responses.

```{r}
left_join(enrollment, survey, by = "id")
```

<br>

### Using a join function to provide a list of students that did not fill out the survey.

```{r}
anti_join(enrollment, survey, by = "id")
```

<br>

### Using a join function to provide a list of students that filled out the survey but have dropped the course.

```{r}
anti_join(survey, enrollment, by = "id")
```

<br>

***

## Case Study 2: Grocery Sales

### Joining the `purchases` and `prices` datasets to obtain a list purchases with their prices. 

```{r}
inner_join(purchases, prices, by = "item")
```

<br>

### Using the join functions and the `summarize` function to obtain an amount of total revenue for purchases. 

```{r}
total <- purchases %>%
  inner_join(prices, by = "item") %>%
  summarize(total_revenue = sum(price))

total
```

<br>

### The total revenue per customer

```{r}
revenuePerCustomer <- purchases %>%
  inner_join(prices, by = "item") %>%
  group_by(customer_id) %>%
  summarize(total_revenue = sum(price))

revenuePerCustomer
```

