---
title: "Lab 12 - Data Import"
author: "Utkal Sirikonda"
output:   
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*** 

<br>

### Download the file `nobel.csv` from the Box folder and load the appropriate packages. 

Make sure to save `nobel.csv` in your same directory (folder) as your Lab12_Template.Rmd file

```{r}
library(tidyverse)
library(readxl)
```

<br>

### Read in the file `nobel.csv` 

```{r}
nobel <- read_csv(file="nobel.csv")
nobel
```

<br>

### Split the data into two objects (STEM and non-STEM)

* Create a new data frame, `nobel_stem`, that filters for the STEM fields (Physics, Medicine, Chemistry, and Economics). 

* Create another data frame, `nobel_nonstem`, that filters for the remaining fields.  

__Hint:__ Use the %in% operator when `filter()`ing

```{r}
stem_categories <- c('Physics', 'Medicine', 'Chemistry', 'Economics')

nobel_stem <- nobel %>%
  filter(category %in% stem_categories)
nobel_stem

nobel_nonstem <- nobel %>%
  filter(!(category %in% stem_categories))
nobel_nonstem
```

<br>

### Write out the two dataframes to `nobel-stem.csv` and `nobel-nonstem.csv`, respectively, to your folder. 

```{r}
nobel_stem
write_csv(nobel_stem, file = "nobel-stem.csv")
write_csv(nobel_nonstem, file = "nobel-nonstem.csv")
```

<br>

### Read in the Excel file called `sales.xlsx` from the Box folder such that is looks like the output in the lab. 


```{r}
sales <- read_excel('sales.xlsx', col_names = FALSE, na = "NA")
colnames(sales) <- c("id", "n")
sales <- sales %>%
  filter(!is.na(n))
sales
```

<br>

### Manipulate the sales data such that it looks like output in lab. 

<br>

```{r}
sales %>%
  mutate(brand = ifelse(str_detect(id, "^Brand"), id, NA)) %>%
  fill(brand) %>% 
  filter(!str_detect(id, "^Brand")) %>% 
  mutate(id = as.numeric(id), n = as.numeric(n)) %>%
  select(brand, id, n) 
```


