---
title: "Lab 4 - Exploratory Data Analysis"
author: ""
output: html_document
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Loading packages
library(tidyverse)
library(palmerpenguins)
```

# EPA 1 - Penguins

## Data

The data we are going to use for this Exploratory Data Analysis is __Palmer Penguins__.  This data set comes from the `palmerpenguins` package in R. 

<br>

### Description:  

Size measurements, clutch observations, and blood isotope ratios for adult foraging Adélie, Chinstrap, and Gentoo penguins observed on islands in the Palmer Archipelago near Palmer Station, Antarctica. Data were collected and made available by Dr. Kristen Gorman and the Palmer Station Long Term Ecological Research (LTER) Program.

<br>


```{r}
# write code to display first 10 rows of the penguins dataset


head(penguins, n=10)

# showing a glimpse of the penguins dataset
# notice I do not have to specify the package since the palmerpenguins and dplyr packages are loaded above
glimpse(penguins)
```

<br>

### The number of observations in the __Palmer Penguins__ dataset is:   

```{r}
# Number of observations in penguins dataset
nrow(penguins)

```

<br>

### The number of variables in the __Palmer Penguins__ dataset is:  

```{r}
# Number of variables in penguins dataset
ncol(penguins)

```

<br>

### Variables 

A list of the variables in the `penguins` dataset is provided below: 

```{r echo=FALSE}
knitr::kable(names(penguins), col.names = "Variables")
```

The image below displays the type of each of these variables. 


```{r}
# Create view of variable types from penguins dataset
library(visdat)
vis_dat(penguins)

```

### Descriptive Statistics

Using basic functions in R, we will explore some simple descriptive statistics for the 2 numeric variables (bill_length_mm and bil_depth_mm).  

```{r}
# Descriptive Statistics for bill_length_mm
mean(x = penguins$bill_length_mm, na.rm = TRUE)


# Descriptive Statistics for bill_depth_mm
mean(x = penguins$bill_depth_mm, na.rm = TRUE)

sd(x = penguins$bill_depth_mm, na.rm = TRUE)
min(x = penguins$bill_depth_mm, na.rm = TRUE)
max(x = penguins$bill_depth_mm, na.rm = TRUE)

```

This is a way to obtain descriptive statistics all at once rather than one line of code at a time.  

```{r}
# Code to display a table of descriptive statistics for the bill_lenghth_mm variable using the summarize function from dplyr package. 
# Note an object stats_length is created that stores the table.  

stats_length <- penguins %>%
  summarize(n= n(),
            x_bar = mean(bill_length_mm, na.rm = TRUE), 
            s = sd(bill_length_mm, na.rm=TRUE), 
            min = min(bill_length_mm, na.rm = TRUE), 
            max(bill_length_mm, na.rm = TRUE))
stats_length


# Enter code to display a table of descriptive statistics for the bill_depth_mm variable 
stat_depth <- penguins %>%
  summarize(n= n(),
            x_bar = mean(bill_depth_mm, na.rm = TRUE), 
            s = sd(bill_depth_mm, na.rm=TRUE), 
            min = min(bill_depth_mm, na.rm = TRUE), 
            max(bill_depth_mm, na.rm = TRUE))
stat_depth

```


## Visualization - Bill Length

We are going to look at the steps to creating a plot for bill length.  

We are going to use `ggplot2` to construct this visualization. 

<br>

### Step 1: Establish Data Source

```{r est_data_pen}
# Start with penguins dataframe
library(ggplot2)
ggplot(data=penguins)

```

### Step 2: Map Bill_Length to X-Axis

```{r map_length_pen}
# map the bill_length_mm variable to the x-axis
ggplot(data=penguins, mapping=aes(x=bill_length_mm))

```

### Step 3: Represent observations as Histogram

```{r rep_length_pen}
# rep points has a histogram 
ggplot(data=penguins, mapping=aes(x=bill_length_mm)) +
  geom_histogram()
```

#### What do you notice about the shape of the graph of `bill_length_mm`?

<span style="color: red;"> 
It looks like a bi modal distribution.

</span>

<br>

## Visualization - Bill Depth

```{r hist_depth}
#Create a histogram for bill_depth_mm
ggplot(data=penguins, mapping=aes(x=bill_depth_mm)) +
  geom_histogram()

```

#### How would you describe the shape of the `bill_depth_mm' variable?

<span style="color: red;"> 
It looks like a multi modal distribution.

</span>

## Visualization - Length vs. Depth

### Step 1: Establish Data Source and Map Variables

```{r est_data_map_pen}
# Start with penguins dataframe
# map bill_depth_mm to x axis 
# map bill_length_mm to y axis

ggplot(data=penguins, mapping=aes(x=bill_depth_mm, y=bill_length_mm))

```

### Step 2: Represent each observation with a Point

```{r rep_point_pen}
# Represent observation as points
ggplot(data=penguins, mapping=aes(x=bill_depth_mm, y=bill_length_mm)) + 
  geom_point()

```

### Step 3: Map Species to Color of Each Point

```{r map_color_pen}
# Map Species to the color of each point

ggplot(data=penguins, mapping=aes(x=bill_depth_mm, y=bill_length_mm, color=species)) + 
  geom_point()

```

### Step 4: Add Main Title

```{r main_title_pen}
# Add a Main Title 
ggplot(data=penguins, mapping=aes(x=bill_depth_mm, y=bill_length_mm, color=species)) + 
  geom_point() +
  labs(title="Bill depth and length")
```

### Step 5: Add Sub Title

```{r sub_title_pen}
# Add a Sub Title 
ggplot(data=penguins, mapping=aes(x=bill_depth_mm, y=bill_length_mm, color=species)) + 
  geom_point() +
  labs(title="Bill depth and length", subtitle="Dimensions for Adelie, Chinstrap, and Gentoo Penguins")
```

### Step 6: Add Axis Labels

```{r axis_lab_pen}
# Add Axis Labels
ggplot(data=penguins, mapping=aes(x=bill_depth_mm, y=bill_length_mm, color=species)) + 
  geom_point() +
  labs(title="Bill depth and length", 
       subtitle="Dimensions for Adelie, Chinstrap, and Gentoo Penguins", 
       x = "Bill depth (mm)", y = "Bill length (mm)")


```


### Step 7: Add Caption

```{r caption}
# Add Caption
ggplot(data=penguins, mapping=aes(x=bill_depth_mm, y=bill_length_mm, color=species)) + 
  geom_point() +
  labs(title="Bill depth and length", 
       subtitle="Dimensions for Adelie, Chinstrap, and Gentoo Penguins", 
       x = "Bill depth (mm)", y = "Bill length (mm)", 
       caption = "Source: Palmer Station LTER / palmerpenguins package")

```

### Step 8: Change Color Scale

```{r color}
# Use a discrete color scale that is designed to be perceived by viewers with common forms of color blindness

ggplot(data=penguins, mapping=aes(x=bill_depth_mm, y=bill_length_mm, color=species)) + 
  geom_point() +
  labs(title="Bill depth and length", 
       subtitle="Dimensions for Adelie, Chinstrap, and Gentoo Penguins", 
       x = "Bill depth (mm)", y = "Bill length (mm)", 
       caption = "Source: Palmer Station LTER / palmerpenguins package") +
  scale_color_viridis_d()

```


