---
title: "Lab6_Template"
author: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source="bg-success")
```

### Setup

```{r message=FALSE, warning=FALSE}
# Load packages needed
library(tidyverse)
library(openintro)
```


```{r}
# Create condensed dataframe loans
glimpse(loans_full_schema)

loans <- loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade, 
         state, annual_income, homeownership, debt_to_income)

```

<br>

#### Using the `loans` dataframe create a histogram for `interest_rate`.  The histogram should have an appropriate amount of bins.  Be sure to provide an apprppriate title and axis labels.

```{r}
ggplot(loans, aes(x = interest_rate)) +
  geom_histogram(binwidth = 1) +
  labs(
    x = "Interest Rates",
    y = "Frequency",
    title = "Interest Rates"
  )

```


#### Describe the shape of the distribution of `interest_rate`. 
The graph looks right skewed.

<br>

#### Using the `loans` dataframe, create a facet histogram of `interest_rate` broken down by `grade`.

```{r}
ggplot(loans, aes(x = interest_rate, fill = grade)) +
  geom_histogram(binwidth = 1, alpha=0.5 ) +
  labs(
    x = "Interest Rate",
    y = "Frequency",
    title = "Interest Rates for Lending Club loans") +
   facet_wrap(~ grade, nrow=7)
```

<br>

#### Using the `loans` dataframe, create a density plot of `interest_rate` separated by `grade`.  Use an appropriate adjustment value.  Use appropriate titles and labels.  


```{r}
ggplot(loans, aes(x = interest_rate, 
                  fill = grade)) +
  geom_density(adjust = 2, 
               alpha = 0.5) +
  labs(
    x = "Interest Rate (%)",
    y = "Density",
    title = "Interest Rates of Lending Club loans", 
    fill = "grade"
  )

```

#### Do you notice anything unusual about this graph?  Explain. 
We can observe that the grades are related to interest rates of the load. And loans with G grade are very dense (interest rate >  30%). Rest of all graded loan rates are distributed normally.
<br>

#### Using the loans dataframe, create boxplots of `interest_rate` sepearted by `grade`, coloring each box a different color based on `grade`.  Use appropriate labels and titles. 

```{r}
ggplot(loans, aes(x = interest_rate,
                  y = grade, fill=grade)) +
  geom_boxplot() +
  labs(
    x = "Interest Rates",
    y = "Grades",
    title = "Interest Rates", 
    subtitle = "by grade of loan"
  )
```


<br>

#### Using the loans dataframe and the idea that you can filter within the ggplot function, create a hexplot of `loan_amount` vs. `interest_rate` for `grade` (Hint: use the `fill` argument) values A thru F (exclude grade G).  Please provide appropriate title(s) and labels. 

```{r}
ggplot(loans %>% filter(grade != "G"), aes(x=loan_amount, y=interest_rate, fill=grade)) +
  geom_hex() +
  labs(title = "Interest Rates vs Loan Amount",
       x="Loan Amount",
       y="Interest Rate")
```

<br>


#### Using the `loans` dataframe find the min, med, max, IQR, and num. of observations for `interest_rate` by `grade`. Label your statistics appropriately.

```{r}
loans %>%
  group_by(grade) %>%
  summarise(
    min_interest = min(interest_rate, na.rm = TRUE),
    med_interest = median(interest_rate, na.rm = TRUE),
    max_interest = max(interest_rate, na.rm = TRUE),
    IQR_interest = IQR(interest_rate, na.rm = TRUE),
    observations = n()
  )
```


<br>

#### Using the `loans` dataframe, create a segmented bar chart of `state` segmented by `homeownership`.  The bars should be horizontal (not vertical) and the length of the bars should be the count for each state.  The graph should have appropriate titles, and labels.

```{r}
ggplot(loans, aes( y = state, fill = homeownership)) +
  geom_bar(position = "fill") +
  labs(
    title = "Loan Counts by State and Homeownership",
    x = "Count",
    y = "State",
    fill = "Homeownership"
  ) 
```

<br>

#### Using the `loans` dataframe, create a frequency table of `grade`.

```{r}
loans %>% group_by(grade) %>% summarise(n = n())
```






