---
title: "HW 5 - Data Wrangling"
name: Utkal Sirikonda
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, prompt = TRUE, comment="")
```

## Set up

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(datasets)
```

1. Reshape `us_rent_income` (from the tidyr package) so that it has one line per state, and two new columns named estimate_income and estimate_rent that contain values from estimate.


```{r q1}

us_rent_income <- us_rent_income %>% 
  select(-moe) %>%
  pivot_wider(
  names_from = variable, 
  values_from = estimate
  ) %>%
  rename("estimate_income" = income, "estimate_rent" = rent)

us_rent_income
```


```{r}
# keep this r chunk as is
s77 <- as_tibble(state.x77, rownames = "State")
```

2. Merge/Join your result from the previous step with `s77` (from the datasets package), and keep all rows. Then, drop rows where any values are missing. You can do this in one or two steps. 

```{r q2}
glimpse(s77)
joint <- full_join(us_rent_income, s77, by=c("NAME"="State")) %>%
  drop_na()

joint
```

3. Add a column containing `state.division`.  Keep only the columns __Name, estimate_income, estimate_rent, Population, Income, Area, Frost, and state.division__  

```{r q3}
joint <- joint %>%
  select(NAME, estimate_income, estimate_rent, Population, Income, Area, Frost) %>%
  mutate('state.division' = state.division)

joint
```

4. Add a column with the proportion of income spent on rent (estimate_rent / estimate_income).

```{r q4}
jointWithPropIncome <- joint %>% 
  mutate(propIncome = round((estimate_rent / estimate_income), digits = 4))
jointWithPropIncome
```

5. Drop rows where Area is not greater than ten times Frost.

```{r q5}
jointWithPropIncome <- jointWithPropIncome %>%
  filter(Area >= 10 * Frost)
jointWithPropIncome
```

6. Summarize the data finding the population-weighted mean rent by geographic division. Sort this summary in descending order by Weighted mean. 

```{r q7}
jointWithPropIncome %>%
  group_by(state.division) %>%
  summarise(MeanRent =round(weighted.mean( estimate_rent, Population), digits = 0)) %>%
  arrange(-MeanRent)
```




