---
title: "User Car Analysis"
author: "Utkal Sirikonda"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: yeti
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
In this project, we study used car data. Analyzing used cars for their reliability is a huge deal for people who are willing to buy one.

## Data Collection

* The dataset is obtained from Kaggle: [used-car-catalog](https://www.kaggle.com/datasets/lepchenkov/usedcarscatalog?resource=download)
* As mentioned in the link, the data is collected from various web resources in order to explore the used cars market and try to build a model that effectively predicts the price of the car based on its parameters (both numerical and categorical).
* This dataset contains used car ads from a popular online catalogs in Belarus on December 2nd, 2019.
* The dataset contains 30 variables, where each row is a car ad, described by manufacturer, model, mileage, production year, and many other properties.


## Importing Data

```{r, echo=FALSE}
library(tidyverse)
library(DT)
```

```{r, warning=FALSE}
input_data <- read.csv("cars.csv", header = TRUE, stringsAsFactors = FALSE)

# datatable(input_data, options = list(pageLength = 10))

glimpse(input_data)
```

## Data Pre-processing and Manipulation

### Some key observations

* Any record without manufacturer_name, model_name, year_produced doesn't describe the car properly, so need to eliminate any such records

* Boolean variables (engine_has_gas, has_warranty, is_exchangeable, feature_*) are stored as "False"/"True" (character strings) instead of actual TRUE/FALSE (logical type).

* Feature variables (feature_0 to feature_9) don't contribute semantically to our analysis as they are not described properly. So, we can either choose to neglect them, or we can just count the number of features and make a new variable out of it.

* Select variables of interest and neglect the rest

* Naming variables can be improvised

### Pre-processing and manipulation:

1. Filter NA records.
```{r}
processed_data <- input_data %>%
  drop_na(manufacturer_name, model_name, year_produced)
```

2. Convert Boolean Columns to Logical Type (TRUE/FALSE).

```{r}
processed_data <- processed_data %>%
  mutate(across(c(engine_has_gas, has_warranty, is_exchangeable,starts_with("feature_")), 
                ~ . == "True"))
```

3. Reshape feature_0 to feature_9 into a feature_count.
```{r}
processed_data <- processed_data %>%
  mutate(feature_count = rowSums(select(., starts_with("feature_"))))
```

4. Filter out unrequired variables.
```{r}
processed_data <- processed_data %>%
  select(manufacturer_name, model_name, year_produced, 
         engine_type, engine_capacity, transmission, 
         body_type, drivetrain, color, mileage = odometer_value, engine_fuel,state,location_region,number_of_photos,
         price_usd, has_warranty, is_exchangeable,duration_listed,up_counter, feature_count)
```


## Processed data
```{r, warning=FALSE}
datatable(processed_data, options = list(pageLength = 10))
```