---
title: "Lab7_Template"
author: "Utkal Sirikonda"
output: html_document
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
```


### Exercise 1

```{r ex1}
table4a
pivot_longer(table4a, !country, names_to = "year", values_to = "cases")
```

<br>


### Exercise 2

```{r ex2}
table2
pivot_wider(table2, names_from = type, values_from = count)
```

<br>

### Exercise 3

```{r ex3}
table3
table3 %>%
  separate(col=rate, into = c("cases", "population"))
```

<br>

*** 


## Case Study

Use the `pivot_longer()` function to produce create a new dataset `who1` 
```{r who1}

who1 <- who %>%
  pivot_longer(
    cols = new_sp_m014:newrel_f65,
    names_to = "key",
    names_transform = list(key = as.character),
    values_to = "cases",
    values_drop_na = TRUE,
  )

who1

who1 %>% 
  count(key)

```

```{r who2, echo=FALSE}
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
who2
```

Use `separate()` to split the `key` variable at each underscore the results should be stored in a dataset named `who3`

```{r who3}
who3 <- who2 %>%
  separate(col=key, into=c("new", "type", "sexage"), sep="_")

who3

who3 %>% 
  count(new)

```

```{r who4, echo=FALSE}
who4 <- who3 %>% 
  select(-new, -iso2, -iso3)
who4
```

Use the separate function to separate `sexage` into `sex` and `age` by splitting after the first character.  Save this dataframe into  new object called `who5`.

```{r who5}
who5 <- who4 %>%
  separate(
    col=sexage, 
    sep = 1,
    into = c("sex", "age")
  )

who5
```














